---
title: "Contour check for *Triadica sebifera*"
---

The goals of this notebook are:
1. To check the original contour data of *Triadica sebifera*.
2. To check contours reconstructed from oriented true EFD normalization for all samples.

Output figures of this notebook are not included in the manuscript, but they are saved in the `results/` directory.

## Setup

```{r}
#| label: setup
library(tools)
library(Momocs)
source("R/normalization.R")
source("R/compare_contours.R")
# color palette for colorblind-friendly visualization
palette("Okabe-Ito")
# set the data directory from environment variable (.Renviron)
data_dir <- Sys.getenv("PROJECT_DATA_DIR")
```

## Load contour data

Load the contour data from the CSV files in the `data/contour_triadica_sebifera/` directory.
There are 30 samples.

```{r}
#| label: load_contours
# file paths of contour CSV files
file_paths_xy <- list.files(
  file.path(data_dir, "data/contour_triadica_sebifera"),
  full.names = TRUE,
  pattern = "\\.csv$"
)
# extract sample IDs from file names
id_name <- file_path_sans_ext(basename(file_paths_xy))

# load contour data into a list of matrices
xy_list <- lapply(file_paths_xy, function(fp) {
  #cat("File:", fp, "\n")
  df <- read.csv(fp)
  df <- df[, c("x", "y")]
  df <- as.matrix(df)
  storage.mode(df) <- "double"
  # if the contour is not closed, add the first point to the end to close it
  if (!all(df[1, ] == df[nrow(df), ])) {
    df <- rbind(df, df[1, ])
  }
  return(df)
})
# assign names to the list of contours
names(xy_list) <- id_name
```

## Load oriented true normalized EFD coefficients

Load the oriented true normalized EFD coefficients from the CSV files in the `data/coefficients_efd_normalized_triadica_sebifera/` directory.
These data are generated with LeafContourEFD.

```{r} 
#| label: load_coefficients
# file paths of EFD coefficient CSV files
file_paths <- list.files(
  file.path(data_dir, "data/coefficients_efd_normalized_triadica_sebifera"),
  full.names = TRUE,
  pattern = "\\.csv$"
)

# load EFD coefficients into a list of data frames
ef_list <- lapply(file_paths, function(fp) {
  #cat("File:", fp, "\n")
  df <- read.csv(fp)
  return(df)
})
```

## Visualize the original contours and the contours reconstructed from oriented true EFD normalization

Overlay the original contour and the reconstructed contour from Oriented True EFD normalization.

```{r}
#| label: compare_contours
layout(matrix(seq_along(xy_list), nrow = 5, byrow = TRUE))
for (i in seq_along(xy_list)) {
  compare_contour_oriented_true_EFD_normalization(
    original = xy_list[[i]],
    ef_normalized = ef_list[[i]],
    label = i,
    nb.h = 5,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5
  )
}
```

```{r}
#| label: save_comparison_plots
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE)
save_file_name <- "supplementary_compare_contours_oriented_true_efd_normalized_triadica_sebifera.svg"
svg(file.path(save_dir, save_file_name), bg = "transparent")
layout(matrix(seq_along(xy_list), nrow = 5, byrow = TRUE))
for (i in seq_along(xy_list)) {
  compare_contour_oriented_true_EFD_normalization(
    original = xy_list[[i]],
    ef_normalized = ef_list[[i]],
    label = i,
    nb.h = 5,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5
  )
}
dev.off()
```
