---
title: "Direction check for *Triadica sebifera*"
---

This notebook compares the original contour and the reconstructed contour from oriented true EFD normalization and true EFD normalization for *Triadica sebifera*.

Output figures in this notebook are used in **Figure 3** of the manuscript.

## Setup

```{r}
#| label: setup
library(tools)
library(Momocs)
source("R/geometry.R")
source("R/normalization.R")
source("R/compare_contours.R")
palette("Okabe-Ito")

# set the data directory from environment variable (.Renviron)
data_dir <- Sys.getenv("PROJECT_DATA_DIR")
```

## Load contour data

Load the contour data from the CSV files in the `data/contour_triadica_sebifera/` directory.

```{r}
#| label: load_contours
file_paths_xy <- list.files(
  file.path(data_dir, "data/contour_triadica_sebifera"),
  full.names = TRUE,
  pattern = "\\.csv$"
)
id_name <- file_path_sans_ext(basename(file_paths_xy))

xy_list <- lapply(file_paths_xy, function(fp) {
  #cat("File:", fp, "\n")
  df <- read.csv(fp)
  df <- df[, c("x", "y")]
  df <- as.matrix(df)
  storage.mode(df) <- "double"
  # if the contour is not closed, add the first point to the end to close it
  if (!all(df[1, ] == df[nrow(df), ])) {
    df <- rbind(df, df[1, ])
  }
  return(df)
})
names(xy_list) <- id_name
```

## Load oriented true normalized EFD coefficients

these data are obtained by LeafContourEFD.

```{r} 
file_paths <- list.files(
  file.path(data_dir, "data/coefficients_efd_normalized_triadica_sebifera"),
  full.names = TRUE,
  pattern = "\\.csv$"
)
ef_list <- lapply(file_paths, function(fp) {
  #cat("File:", fp, "\n")
  df <- read.csv(fp)
  return(df)
})
```

## Oriented true EFD normalization

Compare the original contour and the reconstructed contour from **oriented true EFD normalization** for the first 9 samples.

```{r}
xy_list_example <- xy_list[1:9]
labels <- letters[seq_along(xy_list_example)]
layout(matrix(seq_along(xy_list_example), nrow = 3, byrow = TRUE))
for (i in seq_along(xy_list_example)) {
  compare_contour_oriented_true_EFD_normalization(
    original = xy_list[[i]],
    ef_normalized = ef_list[[i]],
    nb.h = 5,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5,
    label = labels[i]
  )
}
```

```{r}
#| label: reconstruct_contour_oriented_true_EFD_normalization
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE)
save_file_name <- "fig3a_compare_contours_oriented_true_efd_normalized_triadica_sebifera.svg"
svg(file.path(save_dir, save_file_name), bg = "transparent")
xy_list_example <- xy_list[1:9]
labels <- letters[seq_along(xy_list_example)]
layout(matrix(seq_along(xy_list_example), nrow = 3, byrow = TRUE))
for (i in seq_along(xy_list_example)) {
  compare_contour_oriented_true_EFD_normalization(
    original = xy_list[[i]],
    ef_normalized = ef_list[[i]],
    nb.h = 5,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5,
    label = labels[i]
  )
}
dev.off()
```

## True EFD normalization

To assess whether the contours reconstructed agter true EFD normalization are oriented consistently with the original contours, the original contours were randomly rotated prior to applyin true EFD normalization.

```{r}
#| label: rotate_contours
set.seed(123) # set.seed for reproducibility
# rotation angles in degrees
angles <- sample(0:360, length(xy_list), replace = TRUE)
# rotate contours
xy_list_rotated <- mapply(
  function(mat, ang) rotate_xy_centered(mat, ang),
  xy_list,
  angles,
  SIMPLIFY = FALSE
)
```

Compare the original contour and the reconstructed contour from **true EFD normalization** for the first 9 samples.

```{r}
xy_list_rotated_example <- xy_list_rotated[1:9]
labels <- letters[seq_along(xy_list_rotated_example)]
layout(matrix(seq_along(xy_list_rotated_example), nrow = 3, byrow = TRUE))
for (i in seq_along(xy_list_rotated_example)) {
  compare_contour_true_EFD_normalization(
    original = xy_list_rotated_example[[i]],
    nb.h = 5,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5,
    label = labels[i]
  )
}
```

```{r}
#| label: reconstruct_contour_true_EFD_normalization
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE)
save_file_name <- "fig3b_compare_contours_true_efd_normalized_triadica_sebifera.svg"
svg(file.path(save_dir, save_file_name), bg = "transparent")
xy_list_rotated_example <- xy_list_rotated[1:9]
labels <- letters[seq_along(xy_list_rotated_example)]
layout(matrix(seq_along(xy_list_rotated_example), nrow = 3, byrow = TRUE))
for (i in seq_along(xy_list_rotated_example)) {
  compare_contour_true_EFD_normalization(
    original = xy_list_rotated_example[[i]],
    nb.h = 5,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5,
    label = labels[i]
  )
}
dev.off()
```

## Supplementary figures

Visualize all the original contours and the contours reconstructed from oriented true EFD normalization for all samples, and save the plots as SVG files.

```{r}
#| label: reconstruct_contour_oriented_true_EFD_normalization_all
layout(matrix(seq_along(xy_list_rotated), nrow = 5, byrow = TRUE))
for (i in seq_along(xy_list_rotated)) {
  compare_contour_true_EFD_normalization(
    original = xy_list_rotated[[i]],
    nb.h = 5,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5,
    label = i
  )
}
```

```{r}
#| label: save_comparison_plots_all
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE)
save_file_name <- "supplementary_compare_contours_true_efd_normalized_all_triadica_sebifera.svg"
svg(file.path(save_dir, save_file_name), bg = "transparent")
layout(matrix(seq_along(xy_list_rotated), nrow = 5, byrow = TRUE))
for (i in seq_along(xy_list_rotated)) {
  compare_contour_true_EFD_normalization(
    original = xy_list_rotated[[i]],
    nb.h = 5,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5,
    label = i
  )
}
dev.off()
```
