---
title: "Direction check for *Quercus serrata* and *Quercus crispula*"
---

This notebook compares the original contour and the reconstructed contour from oriented true EFD normalization and true EFD normalization for *Quercus serrata* and *Quercus crispula*.

Output figures in this notebook are used in **Figure 2** of the manuscript.

## Setup

```{r}
#| label: setup
library(tools)
library(Momocs)
source("R/geometry.R")
source("R/normalization.R")
source("R/compare_contours.R")
# color palette for colorblind-friendly visualization
palette("Okabe-Ito")
# set the data directory from environment variable (.Renviron)
data_dir <- Sys.getenv("PROJECT_DATA_DIR")
```

## Load species names

Load the species names from the CSV file.

This file contains three columns: `id`, `konara`, and `mizunara`. 
The `id` column corresponds to the sample ID, while the `konara` and `mizunara` columns contain binary values indicating the species classification. 
We will determine the species name based on which column has a value of 1 for each sample.

::: {.callout-note}
## Species name

`konara` means Quercus serrata, and `mizunara` means Quercus crispula.
`konara` (コナラ) represents the standard Japanese name of Quercus serrata, while `mizunara` (ミズナラ) is the standard Japanese name of Quercus crispula.
:::

```{r}
#| label: species_name
df_sp_name <- read.csv(file.path(data_dir, "data/species_name.csv"))
df_sp_name$species <- ifelse(
  df_sp_name$konara < df_sp_name$mizunara,
  "mizunara", # Quercus crispula
  "konara" # Quercus serrata
)
```

## Load contour data

Load the contour data from the CSV files in the `data/contour/` directory.

```{r}
#| label: load_contours
file_paths_xy <- list.files(
  file.path(data_dir, "data/contour_quercus/"),
  full.names = TRUE,
  pattern = "\\.csv$"
)
id_name <- file_path_sans_ext(basename(file_paths_xy))
id_name_head <- lapply(id_name, function(x) {
  parts <- unlist(strsplit(x, "_"))
  return(parts[1])
})
id_name_head <- unlist(id_name_head)
# match species names to contour data
sp_name <- df_sp_name$species[match(id_name_head, df_sp_name$id)]

# load contour data into a list of matrices
xy_list <- lapply(file_paths_xy, function(fp) {
  #cat("File:", fp, "\n")
  df <- read.csv(fp)
  df <- df[, c("x", "y")]
  df <- as.matrix(df)
  storage.mode(df) <- "double"
  # If the contour is not closed, add the first point to the end to close it
  if (!all(df[1, ] == df[nrow(df), ])) {
    df <- rbind(df, df[1, ])
  }
  return(df)
})
names(xy_list) <- id_name
```

## Load oriented true normalized EFD coefficients

these data are obtained by LeafContourEFD.

```{r} 
file_paths <- list.files(
  file.path(data_dir, "data/coefficients_efd_normalized_quercus/"),
  full.names = TRUE,
  pattern = "\\.csv$"
)
ef_list <- lapply(file_paths, function(fp) {
  #cat("File:", fp, "\n")
  df <- read.csv(fp)
  return(df)
})
```

## Oriented true EFD normalization

Compare the original contour and the reconstructed contour from **oriented true EFD normalization** for some samples.

Leftward examples: samples with indices 1, 2, 3, 4, 5, 181, 182, 183, and 184.

```{r}
#| label: compare_oriented_true_EFD_normalization_leftward
idx_left <- c(1:5, 181:184)
labels_left <- letters[1:9]

layout(matrix(seq_along(idx_left), nrow = 3, byrow = TRUE))
for (i in seq_along(idx_left)) {
  compare_contour_oriented_true_EFD_normalization(
    original = xy_list[[idx_left[i]]],
    ef_normalized = ef_list[[idx_left[i]]],
    nb.h = 35,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5,
    label = labels_left[i]
  )
}
```

Rightward examples: samples with indices 322, 374, 375, 376, 378, 595, 596, 746, and 372.

```{r}
#| label: compare_oriented_true_EFD_normalization_rightward
idx_right <- c(322, 374, 375, 376, 378, 595, 596, 746, 372)
labels_right <- letters[10:18]

layout(matrix(seq_along(idx_right), nrow = 3, byrow = TRUE))
for (i in seq_along(idx_right)) {
  compare_contour_oriented_true_EFD_normalization(
    original = xy_list[[idx_right[i]]],
    ef_normalized = ef_list[[idx_right[i]]],
    nb.h = 35,
    lwd = 3,
    cex.text = 5,
    label = labels_right[i]
  )
}
```

```{r}
#| label: reconstruct_contour_oriented_true_EFD_normalization_leftward
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE)
save_file_name <- "fig2a_a-i_compare_contours_oriented_true_efd_normalized_quercus_leftward.svg"
svg(file.path(save_dir, save_file_name), bg = "transparent")
idx_left <- c(1:5, 181:184)
labels_left <- letters[1:9]
layout(matrix(seq_along(idx_left), nrow = 3, byrow = TRUE))
for (i in seq_along(idx_left)) {
  compare_contour_oriented_true_EFD_normalization(
    original = xy_list[[idx_left[i]]],
    ef_normalized = ef_list[[idx_left[i]]],
    nb.h = 35,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 5,
    label = labels_left[i]
  )
}
dev.off()
```

```{r}
#| label: reconstruct_contour_oriented_true_EFD_normalization_rightward
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE)
save_file_name <- "fig2a_j-r_compare_contours_oriented_true_efd_normalized_quercus_rightward.svg"
svg(file.path(save_dir, save_file_name), bg = "transparent")
idx_right <- c(322, 374, 375, 376, 378, 595, 596, 746, 372)
labels_right <- letters[10:18]
layout(matrix(seq_along(idx_right), nrow = 3, byrow = TRUE))
for (i in seq_along(idx_right)) {
  compare_contour_oriented_true_EFD_normalization(
    original = xy_list[[idx_right[i]]],
    ef_normalized = ef_list[[idx_right[i]]],
    nb.h = 35,
    lwd = 3,
    cex.text = 5,
    label = labels_right[i]
  )
}
dev.off()
```

## True EFD normalization

To assess whether the contours reconstructed agter true EFD normalization are oriented consistently with the original contours, the original contours were randomly rotated prior to applyin true EFD normalization.

```{r}
#| label: rotate_contours
set.seed(123) # set.seed for reproducibility
angles <- sample(0:360, length(xy_list), replace = TRUE) # rotation angles in degrees
# rotate contours
xy_list_rotated <- mapply(
  function(mat, ang) rotate_xy_centered(mat, ang),
  xy_list,
  angles,
  SIMPLIFY = FALSE
)
```

Compare the original contour and the reconstructed contour from **true EFD normalization** for samples do not match the original orientation (i.e., samples with indices 1, 2, 3, 4, 5, 181, 182, 183, and 184).

```{r}
#| label: compare_true_EFD_normalization_leftward
idx_left <- c(1:5, 181:184)
labels_left <- letters[1:9]
layout(matrix(seq_along(idx_left), nrow = 3, byrow = TRUE))
for (i in seq_along(idx_left)) {
  compare_contour_true_EFD_normalization(
    original = xy_list_rotated[[idx_left[i]]],
    label = labels_left[i],
    nb.h = 35,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 3
  )
}
```

```{r}
#| label: compare_true_EFD_normalization_rightward
idx_right <- c(322, 374, 375, 376, 378, 595, 596, 746, 372)
labels_right <- letters[10:18]

layout(matrix(seq_along(idx_right), nrow = 3, byrow = TRUE))
for (i in seq_along(idx_right)) {
  compare_contour_true_EFD_normalization(
    original = xy_list_rotated[[idx_right[i]]],
    label = labels_right[i],
    nb.h = 35,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 3
  )
}
```

```{r}
#| label: reconstruct_contour_true_EFD_normalization_leftward
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE)
save_file_name <- "fig2b_a-i_compare_contours_true_efd_normalized_quercus_leftward.svg"
svg(file.path(save_dir, save_file_name), bg = "transparent")
idx_left <- c(1:5, 181:184)
labels_left <- letters[1:9]
layout(matrix(seq_along(idx_left), nrow = 3, byrow = TRUE))
for (i in seq_along(idx_left)) {
  compare_contour_true_EFD_normalization(
    original = xy_list_rotated[[idx_left[i]]],
    label = labels_left[i],
    nb.h = 35,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 3
  )
}
dev.off()
```

```{r}
#| label: reconstruct_contour_oriented_true_EFD_normalization_rightward
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE)
save_file_name <- "fig2c_j-r_compare_contours_true_efd_normalized_quercus_rightward.svg"
svg(file.path(save_dir, save_file_name), bg = "transparent")
idx_right <- c(322, 374, 375, 376, 378, 595, 596, 746, 372)
labels_right <- letters[10:18]

layout(matrix(seq_along(idx_right), nrow = 3, byrow = TRUE))
for (i in seq_along(idx_right)) {
  compare_contour_true_EFD_normalization(
    original = xy_list_rotated[[idx_right[i]]],
    label = labels_right[i],
    nb.h = 35,
    mar = c(0, 0, 0, 0),
    lwd = 3,
    cex.text = 3
  )
}
dev.off()
```

## Supplementary figures

Visualize all the original contours and the contours reconstructed from oriented true EFD normalization for all samples, and save the plots as SVG files.

```{r}
#| label: comparisons_true_EFD_normalization_all
list_id <- list(
  1:50,
  51:100,
  101:150,
  151:200,
  201:250,
  251:300,
  301:350,
  351:400,
  401:450,
  451:500,
  501:550,
  551:600,
  601:650,
  651:700,
  701:746
)
for (ids in list_id) {
  ids <- unlist(ids)
  layout(matrix(1:50, nrow = 5, byrow = TRUE))
  for (id in ids) {
    compare_contour_true_EFD_normalization(
      xy_list_rotated[[id]],
      label = names(xy_list_rotated[id]),
      nb.h = 35,
      lwd = 2,
      mar = c(0, 0, 0, 0),
      cex.text = 1
    )
  }
}
```

```{r}
#| label: save_comparisons_true_EFD_normalization_all
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results/supplementary_compare_contours_true_efd_normalized_quercus/"
dir.create(save_dir, recursive = TRUE, showWarnings = FALSE)
list_id <- list(
  1:50,
  51:100,
  101:150,
  151:200,
  201:250,
  251:300,
  301:350,
  351:400,
  401:450,
  451:500,
  501:550,
  551:600,
  601:650,
  651:700,
  701:746
)

for (ids in list_id) {
  ids <- unlist(ids)
  save_path <- paste0(
    save_dir,
    "/compare_contours_true_efd_normalized_quercus_",
    min(ids),
    "_",
    max(ids),
    ".svg"
  )
  svg(save_path, width = 10, height = 5, bg = "transparent")
  layout(matrix(1:50, nrow = 5, byrow = TRUE))
  for (id in ids) {
    compare_contour_true_EFD_normalization(
      original = xy_list_rotated[[id]],
      label = names(xy_list_rotated[id]),
      nb.h = 35,
      lwd = 2,
      mar = c(0, 0, 0, 0),
      cex.text = 1
    )
  }
  dev.off()
}
```
