---
title: "Functions used in the analysis"
---

The purpose of this notebook is to introduce the functions used in this study.

## `R/compare_contours.R`

`R/compare_contours.R` contains two functions to compare the original contour with the reconstructed contour from EFD coefficients. 

The first function `compare_contour_oriented_true_EFD_normalization` compares the original contour with the reconstructed contour from EFD coefficients normalized by the true normalization method. 

The second function `compare_contour_true_EFD_normalization` compares the original contour with the reconstructed contour from EFD coefficients normalized by the true normalization method without orientation alignment.

### `compare_contour_oriented_true_EFD_normalization()`

```{r}
compare_contour_oriented_true_EFD_normalization <- function(
  original,
  ef_normalized,
  label = NULL,
  nb.h = 35,
  mar = c(0, 0, 0, 0),
  cex.text = 2,
  ...
) {
  # Function to compute E* from first-harmonic EFD coefficients
  compute_E_star <- function(ef) {
    A1 <- ef$an[1]
    B1 <- ef$bn[1]
    C1 <- ef$cn[1]
    D1 <- ef$dn[1]
    # Calculate theta
    theta <- 0.5 *
      atan2(
        2 * (A1 * B1 + C1 * D1),
        (A1^2 + C1^2 - B1^2 - D1^2)
      )

    # Calculate a* and c*
    a_star <- A1 * cos(theta) + B1 * sin(theta)
    c_star <- C1 * cos(theta) + D1 * sin(theta)

    # Scale factor (E*)
    E_star <- sqrt(a_star^2 + c_star^2)
    return(E_star)
  }

  center <- coo_centpos(original) # center position
  original_centered <- sweep(original, 2, center, "-") # center the contour
  ef <- efourier(
    original_centered,
    nb.h = nb.h,
    norm = FALSE
  )
  E <- compute_E_star(ef)
  true_ef_coef <- list(
    ao = unique(ef_normalized$ao),
    co = unique(ef_normalized$co),
    an = ef_normalized$an,
    bn = ef_normalized$bn,
    cn = ef_normalized$cn,
    dn = ef_normalized$dn
  )
  # Reconstruct contour from true EFD coefficients
  reconstruct <- efourier_i(true_ef_coef, nb.h = nb.h)
  df_reconstruct <- as.data.frame(reconstruct)
  df_reconstruct <- df_reconstruct * E
  xlim <- range(c(df_reconstruct$x, original_centered[, 1]))
  ylim <- range(c(df_reconstruct$y, original_centered[, 2]))
  par(mar = mar)
  # original centered contour
  plot(
    original_centered,
    type = "l",
    asp = 1,
    col = adjustcolor("black", red = 0.7, green = 0.7, blue = 0.7),
    xlim = xlim,
    ylim = ylim,
    bty = "n",
    xaxt = "n",
    yaxt = "n",
    xlab = "",
    ylab = "",
    ...
  )

  # reconstructed contour
  lines(
    df_reconstruct,
    asp = 1,
    col = adjustcolor("#FF4B00", alpha.f = 0.9),
    ...
  )

  # label
  if (!is.null(label)) {
    text(
      x = mean(par("usr")[1:2]),
      y = mean(par("usr")[3:4]),
      labels = label,
      cex = cex.text,
      col = "black"
    )
  }
}
```

### `compare_contour_true_EFD_normalization()`

```{r}
compare_contour_true_EFD_normalization <- function(
  original,
  label = NULL,
  nb.h = 35,
  mar = c(0, 0, 0, 0),
  cex.text = 2,
  ...
) {
  center <- coo_centpos(original) # center position
  original_centered <- sweep(original, 2, center, "-") # center the contour

  ef <- efourier(
    original_centered,
    nb.h = nb.h,
    norm = FALSE
  )
  true_ef <- true_normalize(ef, output_meta = TRUE)
  true_ef_coef <- true_ef$ef
  true_ef_coef <- list(
    ao = unique(true_ef_coef$ao),
    co = unique(true_ef_coef$co),
    an = true_ef_coef$an,
    bn = true_ef_coef$bn,
    cn = true_ef_coef$cn,
    dn = true_ef_coef$dn
  )
  # Reconstruct contour from true EFD coefficients
  reconstruct <- efourier_i(true_ef_coef, nb.h = nb.h)
  df_reconstruct <- as.data.frame(reconstruct)
  df_reconstruct <- df_reconstruct * true_ef$E

  xlim <- range(c(df_reconstruct$x, original_centered[, 1]))

  par(mar = mar)
  # original centered contour
  plot(
    original_centered,
    type = "l",
    asp = 1,
    col = adjustcolor("black", red = 0.7, green = 0.7, blue = 0.7),
    xlim = xlim,
    bty = "n",
    xaxt = "n",
    yaxt = "n",
    xlab = "",
    ylab = "",
    ...
  )

  # reconstructed contour
  lines(
    df_reconstruct,
    asp = 1,
    col = adjustcolor("#FF4B00", alpha.f = 0.9),
    ...
  )
  if (!is.null(label)) {
    text(
      x = mean(par("usr")[1:2]),
      y = mean(par("usr")[3:4]),
      labels = label,
      cex = cex.text,
      col = "black"
    )
  }
}
```

## `R/geometry.R`

`R/geometry.R` contains three functions to rotate the contour.

### `rotate_xy()`

```{r}
rotate_xy <- function(df, angle_deg, origin = c(0, 0)) {
  theta <- angle_deg * pi / 180

  x <- df[, 1] - origin[1]
  y <- df[, 2] - origin[2]

  x_new <- x * cos(theta) - y * sin(theta)
  y_new <- x * sin(theta) + y * cos(theta)

  data.frame(x = x_new + origin[1], y = y_new + origin[2])
}
```

### `rotate_xy_centered()`

```{r}
rotate_xy_centered <- function(df, angle_deg) {
  origin <- c(mean(df[, 1]), mean(df[, 2]))
  rotate_xy(df, angle_deg, origin)
}
```

### `rotate_xy_custom_origin()`

```{r}
rotate_xy_custom_origin <- function(df, angle_deg, origins) {
  l <- lapply(seq_along(origins), function(i) {
    rotate_xy(df[i, ], angle_deg[i], origins[[i]])
  })
  do.call(rbind, l)
}
```

## `R/normalization.R`

`R/normalization.R` contains two functions to normalize EFD coefficients by the true normalization method.

### `true_normalize_coe()`

```{r}
true_normalize_coe <- function(
  ef_coe,
  EPS = 1e-10,
  y_sy = TRUE,
  x_sy = TRUE,
  skip_rotation = FALSE,
  verbose = FALSE
) {
  ef_mat <- as.matrix(ef_coe)
  storage.mode(ef_mat) <- "double"

  get_H <- function(cn) {
    idx <- grep("^[ABCD][0-9]+$", cn)
    if (!length(idx)) {
      stop("A/B/C/D indexed columns were not found.")
    }
    max(as.integer(sub("^[ABCD]", "", cn[idx])))
  }

  H <- get_H(colnames(ef_mat))
  nmA <- paste0("A", 1:H)
  nmB <- paste0("B", 1:H)
  nmC <- paste0("C", 1:H)
  nmD <- paste0("D", 1:H)

  need <- c(nmA, nmB, nmC, nmD)
  stopifnot(all(need %in% colnames(ef_mat)))

  norm_one <- function(row) {
    ef <- list(
      an = as.numeric(row[nmA]),
      bn = as.numeric(row[nmB]),
      cn = as.numeric(row[nmC]),
      dn = as.numeric(row[nmD]),
      ao = 0,
      co = 0
    )

    ef_true <- true_normalize(
      ef,
      EPS = EPS,
      y_sy = y_sy,
      x_sy = x_sy,
      skip_rotation = skip_rotation
    )

    if (verbose && skip_rotation) {
      message("skip_rotation is TRUE")
    }

    c(
      setNames(ef_true$an, nmA),
      setNames(ef_true$bn, nmB),
      setNames(ef_true$cn, nmC),
      setNames(ef_true$dn, nmD)
    )
  }

  ef_mat_norm <- t(apply(ef_mat, 1, norm_one))

  rownames(ef_mat_norm) <- rownames(ef_mat)
  ef_mat_norm <- ef_mat_norm[, need, drop = FALSE]
  storage.mode(ef_mat_norm) <- "double"
  ef_mat_norm
}
```

### `efourier_true_norm()`

```{r}
efourier_true_norm <- function(
  coo_out,
  nb_h = 35,
  x_sy = TRUE,
  y_sy = TRUE,
  ...
) {
  ef <- efourier(coo_out, nb.h = nb_h, norm = FALSE)
  ef$coe <- true_normalize_coe(ef$coe, y_sy = y_sy, x_sy = x_sy, ...)
  ef
}
```

## `R/pca_plot.R`

`R/pca_plot.R` contains a function to plot PCA results.

```{r}
pca_plot <- function(x, group = "species", palette = NULL, ...) {
  fac_names <- names(x$fac)
  has_group <- !is.null(fac_names) && group %in% fac_names
  grouping_formula <- if (has_group) stats::reformulate(group) else NULL

  if (is.null(palette)) {
    palette <- if (has_group && exists("my_palette", inherits = TRUE)) {
      get("my_palette", inherits = TRUE)
    } else {
      "grey30"
    }
  }

  plt <- plot_PCA(
    x,
    f = grouping_formula,
    palette = palette,
    points = FALSE,
    morphospace = FALSE,
    morphospace_position = "range",
    chull = TRUE, # convex hull
    #chullfilled = TRUE,
    zoom = 1,
    eigen = FALSE,
    box = FALSE,
    axesnames = FALSE,
    axesvar = FALSE,
    ...
  )
  plt <- layer_axes(plt, col = "black", lwd = 1)
  plt <- layer_ticks(plt, col = "black", lwd = 1)
  plt <- layer_points(plt, pch = 19, cex = 1, transp = 0.5)
  plt <- layer_morphospace_PCA(plt, size = 1, col = "black")
  plt <- layer_axesnames(plt, cex = 1.5, name = "PC") # add axes names
  plt <- layer_axesvar(plt, cex = 1.5) # add axes variance
}
```
