---
title: "PCA of *Triadica sebifera* leaf contours"
---

The goals of this notebook are:  
1. To perform PCA using the coefficients obtained by conventional normalization, oriented true EFD normalization and normalized EFD.  
2. To compare PCA results across these three methods.

## Setup & preprocessing

```{r}
#| label: setup
library(tools)
library(Momocs)
source("R/geometry.R")
source("R/normalization.R")
source("R/pca_plot.R")
# color palette for colorblind-friendly visualization
palette("Okabe-Ito")
# set the data directory from environment variable (.Renviron)
data_dir <- Sys.getenv("PROJECT_DATA_DIR")
```

Load contours from the CSV files in the `data/contour_triadica_sebifera/` directory.

```{r}
#| label: load_contours
file_paths_xy <- list.files(
  file.path(data_dir, "data/contour_triadica_sebifera/"),
  full.names = TRUE,
  pattern = "\\.csv$"
)
xy_list <- lapply(file_paths_xy, function(fp) {
  #cat("File:", fp, "\n")
  df <- read.csv(fp)
  df <- df[, c("x", "y")]
  df <- as.matrix(df)
  storage.mode(df) <- "double"
  # Ensure the contour is closed by checking if the first and last points are the same. If not, append the first point to the end of the matrix.
  if (!all(df[1, ] == df[nrow(df), ])) {
    df <- rbind(df, df[1, ])
  }
  return(df)
})
```

## PCA using conventional normalization

PCA using normalized EFD coefficients (conventional method [@kuhl1982]).  
Normalization is performed using [Momocs](https://momx.github.io/Momocs/) [@bonhomme2014].

```{r}
#| label: pca_efd_normalized
nb_h <- 5 # Number of harmonics
out <- Out(xy_list) # convert to Momocs Out object
ef_norm <- efourier(out, nb.h = nb_h, norm = TRUE, start = FALSE) # EFD with normalization
pca_norm <- PCA(ef_norm, center = TRUE, scale. = FALSE) # PCA with Momocs
pca_plot(pca_norm)
```

```{r}
#| label: pca_efd_normalized_save
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE, recursive = TRUE)
save_file_name <- "fig4d_pca_efd_normalized.svg"
save_file_path <- file.path(save_dir, save_file_name)
svg(
  save_file_path,
  bg = "transparent"
)
pca_plot(pca_norm)
dev.off()
```

## True EFD normalization

PCA using coefficients obtained by true EFD normalization.  
Normalization is performed using the `efourier_true_norm()` function implemented in this project (see [`R/normalization.R`](https://github.com/maple60/leaf-contour-efd-analysis/blob/main/R/normalization.R) for details).

```{r}
#| label: pca_true_efd_normalized
nb_h <- 5 # Number of harmonics
out <- Out(xy_list) # convert to Momocs Out object
# oriented true EFD normalization
ef_true_norm <- efourier_true_norm(
  out,
  nb_h = nb_h,
  x_sy = TRUE,
  y_sy = TRUE,
  skip_rotation = FALSE
)
pca_true_norm <- PCA(ef_true_norm, center = TRUE, scale. = FALSE) # PCA with Momocs
pca_plot(pca_true_norm)
```

```{r}
#| label: pca_true_efd_normalized_save
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE, recursive = TRUE)
save_file_name <- "fig4e_pca_true_efd_normalized.svg"
save_file_path <- file.path(save_dir, save_file_name)
svg(
  save_file_path,
  bg = "transparent"
)
pca_plot(pca_true_norm)
dev.off()
```

## PCA using oriented true EFD normalization

Contours obtained from [LeafContourEFD](https://github.com/maple60/leaf-contour-efd) already have standardized starting points and rotation.
Therefore, we apply only the x-axis symmetry correction step from **true EFD normalization**
to implement oriented **true EFD normalization**.

Although the same results can be obtained from the Fourier coefficients exported by
[LeafContourEFD](https://github.com/maple60/leaf-contour-efd), we use [Momocs](https://momx.github.io/Momocs/) here so we can easily visualize reconstructed contour
changes associated with principal component score shifts.

In a future update, we plan to provide a workflow that does not depend on [Momocs](https://momx.github.io/Momocs/).

```{r}
#| label: pca_oriented_true_efd_normalized
nb_h <- 5 # Number of harmonics
out <- Out(xy_list) # convert to Momocs Out object
# oriented true EFD normalization
ef_true_norm <- efourier_true_norm(
  out,
  nb_h = nb_h,
  x_sy = TRUE,
  y_sy = FALSE,
  skip_rotation = TRUE
)
pca_true_norm <- PCA(ef_true_norm, center = TRUE, scale. = FALSE) # PCA with Momocs
pca_plot(pca_true_norm)
```

```{r}
#| label: pca_oriented_true_efd_normalized_save
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE, recursive = TRUE)
save_file_name <- "fig4f_pca_oriented_true_efd_normalized.svg"
save_file_path <- file.path(save_dir, save_file_name)
svg(
  save_file_path,
  bg = "transparent"
)
pca_plot(pca_true_norm)
dev.off()
```

## References

::: {#refs}
:::
