---
title: "PCA of *Quercus serrata* and *Quercus crispula* leaf contours"
---

Purposes of this notebook are:
1. To perform PCA using the coefficients obtained by conventional normalization, oriented true EFD normalization and normalized EFD.
2. To compare the PCA results among these three methods.

Output figures in this notebook are used in **Figure 4A-C** of the manuscript.

## Setup & preprocessing

```{r}
#| label: setup
library(tools)
library(Momocs)
source("R/geometry.R")
source("R/normalization.R")
source("R/pca_plot.R")
# color palette for colorblind-friendly visualization
palette("Okabe-Ito")
# Quercus serrata and Q. crispula colors
my_palette <- pal_manual(c("#FF4B00", "#005AFF"), transp = 0)
# set the data directory from environment variable (.Renviron)
data_dir <- Sys.getenv("PROJECT_DATA_DIR")
```

Load the species names from the CSV file.

This file contains three columns: `id`, `konara`, and `mizunara`. 
The `id` column corresponds to the sample ID, while the `konara` and `mizunara` columns contain binary values indicating the species classification. 
We will determine the species name based on which column has a value of 1 for each sample.

::: {.callout-note}
## Species name

`konara` means Quercus serrata, and `mizunara` means Quercus crispula.
`konara` (コナラ) represents the standard Japanese name of Quercus serrata, while `mizunara` (ミズナラ) is the standard Japanese name of Quercus crispula.
:::

```{r}
#| label: species_name
df_sp_name <- read.csv(file.path(data_dir, "data/species_name.csv"))
df_sp_name$species <- ifelse(
  df_sp_name$konara < df_sp_name$mizunara,
  "Quercus crispula",
  "Quercus serrata"
)
df_sp_name$species <- factor(
  df_sp_name$species,
  levels = c("Quercus serrata", "Quercus crispula")
)
```

Load contours

```{r}
#| label: load_contours
file_paths_xy <- list.files(
  file.path(data_dir, "data/contour_quercus/"),
  full.names = TRUE,
  pattern = "\\.csv$"
)
xy_list <- lapply(file_paths_xy, function(fp) {
  #cat("File:", fp, "\n")
  df <- read.csv(fp)
  df <- df[, c("x", "y")]
  df <- as.matrix(df)
  storage.mode(df) <- "double"
  # 輪郭が閉じていない場合、最初の点を最後に追加して閉じる
  if (!all(df[1, ] == df[nrow(df), ])) {
    df <- rbind(df, df[1, ])
  }
  return(df)
})
names(xy_list) <- file_path_sans_ext(basename(file_paths_xy))
```

Rotate contour

```{r}
#| label: rotate_contours
set.seed(123) # set.seed for reproducibility
angles <- sample(0:360, length(xy_list), replace = TRUE) # rotation angles in degrees
# rotate contours
xy_list_rotated <- mapply(
  function(mat, ang) rotate_xy_centered(mat, ang),
  xy_list,
  angles,
  SIMPLIFY = FALSE
)
```

To evaluate the normalization methods without bias from the contour’s initial point, we randomize the starting point of each contour. 

This is necessary because the contours obtained with LeafContourEFD already have a standardized starting point. 

To randomize the starting point, we reorder the points in each contour by selecting a starting point at random and rearranging the sequence accordingly.

```{r}
#| label: randomize_starting_point
set.seed(123)
xy_list_rotated_reordered <- lapply(xy_list_rotated, function(xy_rotated) {
  # rotation angles in degrees
  idx_start <- sample(seq_len(nrow(xy_rotated)), 1)
  xy_rotated_reordered <- xy_rotated[
    c(idx_start:nrow(xy_rotated), 1:(idx_start - 1)),
  ]
  return(xy_rotated_reordered)
})
```

Species names corresponding to individual IDs

```{r}
#| label: prepare_ids_species
id_name <- file_path_sans_ext(basename(file_paths_xy)) # IDs with suffixes
# individual IDs without suffixes
id_name_head <- lapply(id_name, function(x) {
  parts <- unlist(strsplit(x, "_"))
  return(parts[1])
})
id_name_head <- unlist(id_name_head)
# species names corresponding to individual IDs
sp_name <- df_sp_name$species[match(id_name_head, df_sp_name$id)]
```

## PCA using conventional normalization

PCA using normalized EFD coefficients (conventional method)
Momocs側で正規化を行います。

```{r}
#| label: pca_efd_normalized
nb_h <- 35 # Number of harmonics
out <- Out(
  xy_list_rotated_reordered,
  fac = data.frame(species = factor(sp_name))
) # convert to Momocs Out object
# EFD with normalization
ef_norm <- efourier(out, nb.h = nb_h, norm = TRUE, start = FALSE)
pca_norm <- PCA(ef_norm, center = TRUE, scale. = FALSE) # PCA with Momocs

idx_max <- which.max(pca_norm$x[, 2])
idx_min <- which.min(pca_norm$x[, 2])

pca_plot(pca_norm) # visualize PCA results
```

```{r}
#| label: pca_efd_normalized_save
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE, recursive = TRUE)
save_file_name <- "fig4a_pca_efd_normalized_quercus.svg"
save_file_path <- file.path(save_dir, save_file_name)
svg(
  save_file_path,
  bg = "transparent"
)
pca_plot(pca_norm)
dev.off()
```

## PCA using true EFD normalization

```{r}
#| label: pca_true_efd_normalized
nb_h <- 35 # Number of harmonics
# convert to Momocs Out object
out <- Out(
  xy_list_rotated_reordered,
  fac = data.frame(species = factor(sp_name))
)
# oriented true EFD normalization
ef_true_norm <- efourier_true_norm(
  out,
  nb_h = nb_h,
  x_sy = TRUE,
  y_sy = TRUE,
  skip_rotation = FALSE
)
# PCA with Momocs
pca_true_norm <- PCA(ef_true_norm, center = TRUE, scale. = FALSE)
# flip PC axes for better visualization
pca_true_norm <- flip_PCaxes(pca_true_norm, 1)
pca_plot(pca_true_norm)
```

```{r}
#| label: pca_true_efd_normalized_save
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE, recursive = TRUE)
save_file_name <- "fig4b_pca_true_efd_normalized.svg"
save_file_path <- file.path(save_dir, save_file_name)
svg(
  save_file_path,
  bg = "transparent"
)
pca_plot(pca_true_norm)
dev.off()
```

## PCA using oriented true EFD normalization

(後で英語に直す)
LeafContourEFDで得られた輪郭はすでに開始点、回転が標準化されています。
この輪郭にtrue EFD noramlizationのx対照補正のみを行うことで、oriente true EFD normalizationを実現します。

LeafContourEFDで得られたフーリエ係数を使用しても同じ結果が得られますが、今回はMomocsの関数を利用して、PC値の変化に応じた再構築輪郭の変化を簡単にPCAプロット上に可視化できるようにします。

将来的にはMomocsに依存しない方法も提供する予定です。

```{r}
#| label: pca_oriented_true_efd_normalized
nb_h <- 35 # Number of harmonics
out <- Out(xy_list, fac = data.frame(species = factor(sp_name))) # convert to Momocs Out object
# oriented true EFD normalization
ef_true_norm <- efourier_true_norm(
  out,
  nb_h = nb_h,
  x_sy = TRUE,
  y_sy = FALSE,
  skip_rotation = TRUE
)
# PCA with Momocs
pca_true_norm <- PCA(ef_true_norm, center = TRUE, scale. = FALSE)
```

Visualize PCA results

```{r}
#| label: pca_oriented_true_efd_normalized_plot
pca_plot(pca_true_norm)
```

```{r}
#| label: pca_oriented_true_efd_normalized_save
#| eval: false
#| code-fold: true
#| code-summary: "Save the plot as SVG file"
save_dir <- "results"
dir.create(save_dir, showWarnings = FALSE, recursive = TRUE)
save_file_name <- "fig4c_pca_oriented_true_efd_normalized.svg"
save_file_path <- file.path(save_dir, save_file_name)
svg(
  save_file_path,
  bg = "transparent"
)
pca_plot(pca_true_norm)
dev.off()
```
